### 2025-09-26 site/index.qmd
updated publications

---
title: "Nic Steyn"
image: files/index/me.jpg
page-layout: full

about:
  id: about
  template: trestles
  image-width: 70%
  links:
    - icon: github
      text: Github
      href: https://github.com/nicsteyn2
    - text: "{{< ai google-scholar >}} Google Scholar"
      href: https://scholar.google.com/citations?user=DyNrh4cAAAAJ
      target: _blank
    - icon: envelope
      text: Email
      href: mailto:nic.steyn@outlook.com
    - icon: linkedin
      text: LinkedIn
      href: https://www.linkedin.com/in/nic-steyn/
      target: _blank
    - text: "{{< ai orcid >}} ORCID"   # use academic icons, "" are needed
      href: https://orcid.org/0000-0001-8904-2941
      target: _blank
    # - icon: globe
    #   text: Department of Statistics
    #   href: https://www.stats.ox.ac.uk/people/nicholas-steyn
    #   target: _blank

# listing:
#   - id: recent-posts
#     contents: 
#       - posts
#       - posts/post-links.yml
#     exclude:
#       title: "Participating APPAM 2024"
#     type: grid
#     max-items: 3
# #    image-height: 150px
#     page-size: 3 #set same as max-times to remove the page number setting
#     fields: [image, date, title, author, description]
#     max-description-length: 75
#     sort: 
#       - "date desc"
#       - "title asc"
---

:::{#about}

## Bio

I am a doctoral student on the [Modern Statistics and Statistical Machine Learning CDT](https://statml.io) at the University of Oxford, supervised by Professor Christl Donnelly and Dr Kris V Parag.

My thesis focuses on robust statistical methods for analysing infectious disease dynamics with an emphasis on **uncertainty quantification**. Recent work has included developing a theoretical/philosophical decision-theoretic framework targeted for the practicalities of epidemiology ([see our preprint here](https://doi.org/10.48550/arXiv.2509.20013)), and extending existing popular methods in the field to improve the robustness of their UQ ([see our published paper here](https://doi.org/10.1093/aje/kwaf165)).

Further research from my thesis combines Bayesian inference, [sequential Monte Carlo methods](http://nicsteyn2.github.io/SMCforRt/) ([see also](https://doi.org/10.1002/sim.70204)), and epidemiological modelling, with applications to standard epidemiological data, as well as [survey data](https://doi.org/10.1101/2025.04.16.25325936) ([see also]((https://doi.org/10.1101/2025.03.03.25323250))) and [wastewater-based surveillance](https://github.com/nicsteyn2/NZWastewaterModelling).

During my studies, I have also worked as a research assistant in the [*The Global Reference Group for Children Affected by Crisis*](https://www.spi.ox.ac.uk/the-global-reference-group-on-children-affected-by-covid-19#collapse3701021), estimating [orphanhood in Brazil](https://doi.org/10.1101/2025.01.31.25321479) during the COVID-19 pandemic, and have held a variety of tutoring positions (departmental, college, and inter-departmental courses).

Past experience includes modelling for policymakers during the COVID-19 pandemic in my home country of New Zealand (March 2020 - September 2021), a quantitative trading internship at Optiver in Sydney (November 2019 - January 2020), and a data analysis internship at a startup in Bangkok (January - February 2018).

## Education

**University of Oxford** | Oxford, United Kingdom<br>
DPhil in Statistics | 2021 - 2025

**University of Canterbury** | Christchurch, New Zealand<br>
BSc (Hons) Compuational and Applied Mathematics | 2019<br>
Awarded with first class

**University of Canterbury** | Christchurch, New Zealand<br>
BSc Financial Engineering and Statistics | 2016 - 2018<br>
GPA: 8.96 out of 9

I also studied statistics at **National University of Singapore** on an exchange programme in 2018.

:::